---
title: "Stats_218_HW2"
author: "Zihan Lin"
date: "2024-10-15"
output: pdf_document
---


```{r warning=FALSE, message=FALSE}
library(networkdata)
library(sna)
library(network)
library(igraph)
library(degreenet)
library(tidyverse)
data(florentine)
data(butland_ppi)
data(addhealth9)
data(tribes)
```

\newpage

## Question 1
### a. 
```{r}
only_boy <- which(addhealth9$X[,"female"] == 0)
health_ntwk <- network(addhealth9$E[addhealth9$E[,1] %in% only_boy 
                                    & addhealth9$E[,2] %in% only_boy,], 
                       directed = FALSE)
largest_health <- component.dist(health_ntwk)$membership == which.max(component.dist(health_ntwk)$csize)
larg_health_ntwk <- network(health_ntwk[largest_health, largest_health], directed = FALSE)

evcent_health <- sna::evcent(larg_health_ntwk, gmode = "graph")
barplot(evcent_health)
summary(evcent_health)

centralization(dat = larg_health_ntwk, FUN = sna::evcent, mode = "graph")
```

\newpage

```{r}
tribes_pos_ntwk <- network(tribes[,,"pos"], directed = FALSE)
largest_tribes <- component.dist(tribes_pos_ntwk)$membership == which.max(component.dist(tribes_pos_ntwk)$csize)

larg_tribes_pos_ntwk <- network(tribes_pos_ntwk[largest_tribes, largest_tribes], directed = FALSE)
evcent_tribes_large <- sna::evcent(larg_tribes_pos_ntwk, gmode = "graph")
barplot(evcent_tribes_large)
summary(evcent_tribes_large)

centralization(dat = larg_tribes_pos_ntwk, FUN = sna::evcent, mode = "graph")
```

\newpage

```{r}
but_ntwk <- network(butland_ppi, directed = FALSE)
larg_but <- component.dist(but_ntwk)$membership == which.max(component.dist(but_ntwk)$csize)

larg_but_ntwk <- network(but_ntwk[larg_but, larg_but], directed = FALSE)
evcent_but_larg <- sna::evcent(larg_but_ntwk, gmode = "graph")
barplot(evcent_but_larg)
summary(evcent_but_larg)

centralization(dat = larg_but_ntwk, FUN = sna::evcent, mode = "graph")
```


\newpage

### b. 
```{r}
close_health <- sna::closeness(larg_health_ntwk, gmode = "graph")
close_tribes <- sna::closeness(larg_tribes_pos_ntwk, gmode = "graph")
close_but <- sna::closeness(larg_but_ntwk, gmode = "graph")

barplot(close_health)
barplot(close_tribes)
barplot(close_but)

summary(close_health)
summary(close_tribes)
summary(close_but)

centralization(dat = larg_health_ntwk, FUN = sna::closeness, mode = "graph")
centralization(dat = larg_tribes_pos_ntwk, FUN = sna::closeness, mode = "graph")
centralization(dat = larg_but_ntwk, FUN = sna::closeness, mode = "graph")
```

\newpage

### c. 
```{r}
between_health <- sna::betweenness(larg_health_ntwk, gmode = "graph")
between_tribes <- sna::betweenness(larg_tribes_pos_ntwk, gmode = "graph")
between_but <- sna::betweenness(larg_but_ntwk, gmode = "graph")

barplot(between_health)
barplot(between_tribes)
barplot(between_but)

summary(between_health)
summary(between_tribes)
summary(between_but)

centralization(dat = larg_health_ntwk, FUN = sna::betweenness, mode = "graph")
centralization(dat = larg_tribes_pos_ntwk, FUN = sna::betweenness, mode = "graph")
centralization(dat = larg_but_ntwk, FUN = sna::betweenness, mode = "graph")
```

\newpage

### d. 
```{r}
degree_health <- sna::degree(larg_health_ntwk, gmode = "graph") %>% table()
degree_tribes <- sna::degree(larg_tribes_pos_ntwk, gmode = "graph") %>% table()
degree_but <- sna::degree(larg_but_ntwk, gmode = "graph") %>% table()

barplot(degree_health)
barplot(degree_tribes)
barplot(degree_but)

summary(degree_health)
summary(degree_tribes)
summary(degree_but)
```

\newpage

```{r}
health_summary <- data.frame(
  Centrality = c("Degree", "Eigenvector", "Closeness", "Betweenness"),
  Min = c(min(degree_health), min(evcent_health), min(close_health), min(between_health)),
  Max = c(max(degree_health), max(evcent_health), max(close_health), max(between_health)),
  Mean = c(mean(degree_health), mean(evcent_health), mean(close_health), mean(between_health)),
  SD = c(sd(degree_health), sd(evcent_health), sd(close_health), sd(between_health))
)

tribes_summary <- data.frame(
  Centrality = c("Degree", "Eigenvector", "Closeness", "Betweenness"),
  Min = c(min(degree_tribes), min(evcent_tribes_large), min(close_tribes), min(between_tribes)),
  Max = c(max(degree_tribes), max(evcent_tribes_large), max(close_tribes), max(between_tribes)),
  Mean = c(mean(degree_tribes), mean(evcent_tribes_large), mean(close_tribes), mean(between_tribes)),
  SD = c(sd(degree_tribes), sd(evcent_tribes_large), sd(close_tribes), sd(between_tribes))
)

but_summary <- data.frame(
  Centrality = c("Degree", "Eigenvector", "Closeness", "Betweenness"),
  Min = c(min(degree_but), min(evcent_but_larg), min(close_but), min(between_but)),
  Max = c(max(degree_but), max(evcent_but_larg), max(close_but), max(between_but)),
  Mean = c(mean(degree_but), mean(evcent_but_larg), mean(close_but), mean(between_but)),
  SD = c(sd(degree_but), sd(evcent_but_larg), sd(close_but), sd(between_but))
)

health_summary
tribes_summary
but_summary
```

These measures capture different aspects of node importance and centrality, so while there may be correlations between them, they are not measuring the same thing. Eigenvector centrality might emphasize the influence of connected nodes, while closeness focuses on proximity, and betweenness highlights the node’s control over information flow.



### e. 
```{r}
deg_cent_health <- centralization(larg_health_ntwk, FUN = sna::degree, mode = "graph")
deg_cent_tribes <- centralization(larg_tribes_pos_ntwk, FUN = sna::degree, mode = "graph")
deg_cent_but <- centralization(larg_but_ntwk, FUN = sna::degree, mode = "graph")

close_cent_health <- centralization(larg_health_ntwk, FUN = sna::closeness, mode = "graph")
close_cent_tribes <- centralization(larg_tribes_pos_ntwk, FUN = sna::closeness, mode = "graph")
close_cent_but <- centralization(larg_but_ntwk, FUN = sna::closeness, mode = "graph")

between_cent_health <- centralization(larg_health_ntwk, FUN = sna::betweenness, mode = "graph")
between_cent_tribes <- centralization(larg_tribes_pos_ntwk, FUN = sna::betweenness, mode = "graph")
between_cent_but <- centralization(larg_but_ntwk, FUN = sna::betweenness, mode = "graph")

evcent_cent_health <- centralization(larg_health_ntwk, FUN = sna::evcent, mode = "graph")
evcent_cent_tribes <- centralization(larg_tribes_pos_ntwk, FUN = sna::evcent, mode = "graph")
evcent_cent_but <- centralization(larg_but_ntwk, FUN = sna::evcent, mode = "graph")

centralization_summary <- data.frame(
  Network = c("ppi", "addhealth", "tribes"),
  Degree = round(c(deg_cent_but, deg_cent_health, deg_cent_tribes),2),
  Closeness = round(c(close_cent_but, close_cent_health, close_cent_tribes),2),
  Betweenness = round(c(between_cent_but, between_cent_health, between_cent_tribes),2),
  Eigenvector = round(c(evcent_cent_but, evcent_cent_health, evcent_cent_tribes),2)
)

centralization_summary
```


```{r echo = FALSE}
health_igraph <- graph_from_adjacency_matrix(as.matrix(as.sociomatrix(larg_health_ntwk)))
tribes_igraph <- graph_from_adjacency_matrix(as.matrix(as.sociomatrix(larg_tribes_pos_ntwk)))
but_igraph <- graph_from_adjacency_matrix(as.matrix(as.sociomatrix(larg_but_ntwk)))

vertex_health <- vertex_connectivity(health_igraph)
vertex_tribes <- vertex_connectivity(tribes_igraph)
vertex_but <- vertex_connectivity(but_igraph)

edge_health <- edge_connectivity(health_igraph)
edge_tribes <- edge_connectivity(tribes_igraph)
edge_but <- edge_connectivity(but_igraph)

vertex_cut_health <- min_cut(health_igraph, value.only = FALSE)$partition1
vertex_cut_tribes <- min_cut(tribes_igraph, value.only = FALSE)$partition1
vertex_cut_but <- min_cut(but_igraph, value.only = FALSE)$partition1

edge_cut_health <- min_cut(health_igraph, value.only = FALSE)$cut
edge_cut_tribes <- min_cut(tribes_igraph, value.only = FALSE)$cut
edge_cut_but <- min_cut(but_igraph, value.only = FALSE)$cut
```



```{r echo = FALSE}
network_summary <- data.frame(
  Network = c("Health", "Tribes", "Butland"),
  Vertex_Connectivity = c(vertex_connectivity(as.undirected(health_igraph)),
                          vertex_connectivity(as.undirected(tribes_igraph)),
                          vertex_connectivity(as.undirected(but_igraph))),
  
  Edge_Connectivity = c(edge_connectivity(as.undirected(health_igraph)),
                        edge_connectivity(as.undirected(tribes_igraph)),
                        edge_connectivity(as.undirected(but_igraph))),
  
  Vertex_Cut = c(paste(min_separators(as.undirected(health_igraph))),
                 paste(min_separators(as.undirected(tribes_igraph))),
                 paste(min_separators(as.undirected(but_igraph)))),
  
  Edge_Cut = c(min_cut(as.undirected(health_igraph)),
               min_cut(as.undirected(tribes_igraph)),
               min_cut(as.undirected(but_igraph)))
)

```

```{r echo = FALSE}
ver_heal <- vertex_connectivity(as.undirected(health_igraph))
ver_trib <- vertex_connectivity(as.undirected(tribes_igraph))
ver_but <- vertex_connectivity(as.undirected(but_igraph))
  
edge_heal <- edge_connectivity(as.undirected(health_igraph))
edge_trib <- edge_connectivity(as.undirected(tribes_igraph))
edge_but <- edge_connectivity(as.undirected(but_igraph))

min_ver_heal <- min_separators(as.undirected(health_igraph))
min_ver_trib <- min_separators(as.undirected(tribes_igraph))
min_ver_but <- min_separators(as.undirected(but_igraph))
  
min_edge_heal <- min_cut(as.undirected(health_igraph), value.only = FALSE)$cut
min_edge_trib <- min_cut(as.undirected(tribes_igraph), value.only = FALSE)$cut
min_edge_but <- min_cut(as.undirected(but_igraph), value.only = FALSE)$cut




```

```{r echo=FALSE}
#"Health Network - Vertex Connectivity:", `r ver_heal`

#"Health Network - Edge Connectivity:", `r edge_heal`

#"Health Network - Vertex Cut:", `r min_ver_heal`

# "Health Network - Edge Cut:", `r min_edge_heal`

#"Tribes Network - Vertex Connectivity:", `r ver_trib`

#"Tribes Network - Edge Connectivity:", `r edge_trib`

#"Tribes Network - Vertex Cut:", `r min_ver_trib`

#"Tribes Network - Edge Cut:", `r min_edge_trib`

#"Butland Network - Vertex Connectivity:", `r ver_but`

#"Butland Network - Edge Connectivity:", `r edge_but`

#"Butland Network - Vertex Cut:", `r min_ver_but`

#"Butland Network - Edge Cut:", `r min_edge_but`
```




\newpage

## Question 2
### a. 
```{r}
flom_ntwk <- network(flomarriage, directed = FALSE)
larg_flom <- component.dist(flom_ntwk)$membership == which.max(component.dist(flom_ntwk)$csize)

larg_flom_ntwk <- network(flom_ntwk[larg_flom, larg_flom], directed = FALSE)
flom_igraph <- graph_from_adjacency_matrix(as.matrix(as.sociomatrix(larg_flom_ntwk)))
```


```{r}
ver_flom <- vertex_connectivity(as.undirected(flom_igraph))
edge_flom <- edge_connectivity(as.undirected(flom_igraph))
min_ver_flom <- min_separators(as.undirected(flom_igraph))
min_edge_flom <- min_cut(as.undirected(flom_igraph), value.only = FALSE)$cut

ver_flom
edge_flom
min_ver_flom
min_edge_flom
```

```{r echo=FALSE}
#"Flomarriage Network - Vertex Connectivity:", `r ver_flom`

#"Flomarriage Network - Edge Connectivity:", `r edge_flom`

#Flomarriage Network - Vertex Cut:", `r min_ver_flom`

#"Flomarriage Network - Edge Cut:", `r min_edge_flom`

```





\newpage

## Question 3
### a. 
```{r}
ccsb <- read.table("CCSB-Y2H.txt")
ccsb <- network(ccsb[, 1:2], directed = TRUE, loops = TRUE)
ccsb[cbind(1:network.size(ccsb), 1:network.size(ccsb))] <- 0
ccsb
plot(ccsb)
```

\newpage

### b. 
```{r}
out_degrees <- sna::degree(ccsb, cmode = "outdegree")
in_degrees <- sna::degree(ccsb, cmode = "indegree")
total_degrees <- sna::degree(ccsb)
```

```{r warning=FALSE}
zipf <- adpmle(total_degrees)
# summary(zipf)
yule <- ayulemle(total_degrees)
# summary(yule)
waring <- awarmle(total_degrees)
# summary(waring)
possion <- apoimle(total_degrees)
# summary(possion)
cmp <- acmpmle(total_degrees)
# summary(cmp)
```



### c. 
```{r}
aic_zipf <- lldpall(zipf$theta, total_degrees)
aic_yule <- llyuleall(yule$theta, total_degrees)
aic_waring <- llwarall(waring$theta, total_degrees)
aic_poisson <- llpoiall(possion$theta, total_degrees)
aic_cmp <- llcmpall(cmp$theta, total_degrees)
```


```{r}
data.frame(zipf = aic_zipf,
           yule = aic_yule,
           waring = aic_waring,
           possion = aic_poisson,
           cmp = aic_cmp)
```

The CMP model fits the data best according to all the measures we looked at, followed by the Waring, Yule, and Zipf models. The Poisson model does a very poor job and doesn't fit the data well at all. The CMP model works well because it can handle some of the complexity in the data that the other models can't, which is why it comes out on top. The other models still do a decent job, but the CMP model is the best choice based on the results.



\newpage

## Question 4
### a. 
```{r}
ccsb <- read.table("CCSB-Y2H.txt")
undirected_ccsb_network <- network(ccsb[, 1:2], directed = FALSE, loops = TRUE)
undirected_ccsb_network[cbind(1:network.size(undirected_ccsb_network), 1:network.size(undirected_ccsb_network))] <- 0
ccsb_comp <- component.dist(undirected_ccsb_network)
comp_size <- ccsb_comp$csize
largest_size <- max(comp_size)

largest_size
length(comp_size)
```

The largest component of the network contains 964 nodes.

Yes, the network likely has a giant component, as the largest component includes 964 out of 1278 nodes.


### b. 
```{r}
larg_ccsb <- component.dist(undirected_ccsb_network)$membership == which.max(component.dist(undirected_ccsb_network)$csize)
larg_ccsb_ntwk <- network(undirected_ccsb_network[larg_ccsb, larg_ccsb], directed = FALSE)
plot(larg_ccsb_ntwk)
```

\newpage


### c. 
```{r}
geodesic_distances <- geodist(undirected_ccsb_network, inf.replace = NA)$gdist
summary(c(geodesic_distances) %>% na.omit())
```


```{r}
N = 1278
L = 964

proportion_reachable = ((L * (L - 1)) / 2) / ((N * (N - 1)) / 2)
proportion_reachable
```


```{r}
mean_geodesic_distance <- mean(geodesic_distances, na.rm = TRUE)
mean_geodesic_distance
```


```{r}
node_degrees <- sna::degree(undirected_ccsb_network)
num_isolates <- sum(node_degrees == 0)
num_isolates
```

















